Parsing Postfix log files

### 2008-11-22 doc/logparser-results.tex
Switched to new glossaries package. A couple of other XXX items have been done too.

Parsing efficiency is an obvious concern when the parser routinely needs to
parse large log files.  The server that generated the log files used
in testing this parser accepts approximately 10,000 mails for 700 users a
day; \graphref{Mails received per day} shows that as expected there are far
more mails received on weekdays than at weekends.   Median log file size is
50MB, containing 285,000 log lines --- large scale mail servers would have
much larger log files.  Note that \graphref{Mails received per day} and
\tableref{Number of mails received per day: statistics} show the number of
mails received by \SMTP{} only; in particular the mail loops noticeable in
other graphs do not contribute to these figures.

@@

    \item Postfix accepts the mail for username@domain, but during delivery
        it notices that the \texttt{Delivered-To} header already contains
        the address it is currently delivering to, and therefore sends a
        bounce notification with sender address \texttt{<>} to the original
        sender: username@domain.

@@

Full coverage of log lines is easy to achieve yet hard to maintain.  It is
easy to achieve full coverage for a limited set of log files (at the time
of writing the parser is tested with \numberOFrules{} rules, fully parsing
\numberOFlogFILESall{} contiguous log files from Postfix 2.2 through to
Postfix 2.5), and new rules are easy to add.  Maintaining full coverage is
hard because other servers have different restrictions with custom
messages, \DNSBL{} messages change over time, major releases of Postfix
change warning messages (usually adding more information), etc.,\ so over
time the log lines drift and change.  \Graphref{rule hits graph} shows the
number of hits for each rule over all \numberOFlogFILES{} log files.  The
number of hits per rule is quite unevenly spread, resembling a Power Law
distribution --- it is obvious that a small number of rules match the vast
majority of the lines, and more than half the rules match fewer than 100
times.

@@

Coverage of mails is much more difficult to determine accurately than
coverage of log lines.  The parser can dump its state tables in a human
readable form; examining these tables with reference to the log files is
the best way to detect mails that were not handled properly (many of the
complications discussed in \sectionref{complications} were
detected in this way).  The parser issues warnings when it detects any
errors, some of which may alert the user to a problem, e.g.\ when a queueid
is reused before the previous mail is fully dealt with, when a queueid or
\pid{} is not found in the state tables,\footnote{There will often be
warnings about a missing queueid or \pid{} in the first few hundred or
thousand log lines because the earlier log lines for those connections or
mails are in the previous log file; loading the saved state from the
previous log file will solve this problem.} or when there are problems
tracking a child mail (see \sectionref{tracking re-injected mail}).  There
should be few or no warnings when parsing, and when finished parsing the
state table should only contain entries for mails that had yet to be
delivered when the log files ended, or were accepted before the log files
began.

At the time of writing the parser is being tested with \numberOFlogFILES{}
log files.  There are 5 warnings produced, but because the parser errs on
the side of producing more warnings rather than fewer, those 5 warnings
represent 3 instances of 1 problem: 3 connections started before the first
log file, so their initial log entries are missing, leading to warnings
when their log lines are parsed.

@@

    \item 9 mails whose initial log entries were not present in the log
        files.

@@

    \item Queueids and \pids{} naturally group together log lines belonging
        to one mail or connection respectively; it is extremely unlikely
        that a log line would be associated with the wrong connection.

@@

It is important that the log lines used are contiguous so that all log
entries are present for as many of the connections as possible.  This log
segment was parsed with all debugging options enabled, resulting in 167,448
lines of output.\footnote{A mean of 27.908 lines of output per line of
input; each connection has 30 debugging lines, plus 21 debugging lines per
result.  Connections which have been cloned will have the cloned connection
in their debugging output, plus another 33 debugging lines.  Those numbers
are approximate, and may vary $\pm{}$ 2.  There is a linear relationship
between the number of log lines and debugging lines: $33(connections) +
30(accepted~~mails) + 21(results)$.  This formula is an approximation only,
and has not been rigorously verified.}  All 167,448 lines were examined in
conjunction with the input log and a dump of the resulting database,
verifying that for each of the input lines the parser used the correct rule
and executed the correct action, that in turn produced the correct result
and inserted the correct data in the database.  The log segment produced 4
warnings, 10 mails remaining in the state tables, and 1625 connections
correctly entered in the database.

>>>>

Parsing efficiency is an obvious concern when the parser routinely needs to
parse large log files.  The server that generated the log files used in
testing this parser accepts approximately 10,000 mails for 700 users a day;
\graphref{Mails received per day} shows that as expected there are far more
mails received on weekdays than at weekends.   Median log file size is
50MB, containing 285,000 log lines --- large scale mail servers would have
much larger log files.  Note that \graphref{Mails received per day} and
\tableref{Number of mails received per day: statistics} show the number of
mails received by \gls{SMTP} only; in particular the mail loops noticeable
in other graphs do not contribute to these figures.

@@

    \item Postfix accepts the mail for username@domain, but during delivery
        it notices that the \texttt{Delivered-To} header already contains
        the address it is currently delivering to, and therefore sends a
        bounce notification with sender address \texttt{<>}\glsadd{<>} to
        the original sender: username@domain.

@@

Full coverage of log lines is easy to achieve yet hard to maintain.  It is
easy to achieve full coverage for a limited set of log files (at the time
of writing the parser is tested with \numberOFrules{} rules, fully parsing
\numberOFlogFILESall{} contiguous log files from Postfix 2.2 through to
Postfix 2.5), and new rules are easy to add.  Maintaining full coverage is
hard because other servers have different restrictions with custom
messages, \gls{DNSBL} messages change over time, major releases of Postfix
change warning messages (usually adding more information), etc.,\ so over
time the log lines drift and change.  \Graphref{rule hits graph} shows the
number of hits for each rule over all \numberOFlogFILES{} log files.  The
number of hits per rule is quite unevenly spread, resembling a Power Law
distribution --- it is obvious that a small number of rules match the vast
majority of the lines, and more than half the rules match fewer than 100
times.

@@

Coverage of mails is much more difficult to determine accurately than
coverage of log lines.  The parser can dump its state tables in a human
readable form; examining these tables with reference to the log files is
the best way to detect mails that were not handled properly (many of the
complications discussed in \sectionref{complications} were
detected in this way).  The parser issues warnings when it detects any
errors, some of which may alert the user to a problem, e.g.\ when a queueid
is reused before the previous mail is fully dealt with, when a queueid or
\gls{pid} is not found in the state tables,\footnote{There will often be
warnings about a missing queueid or \gls{pid} in the first few hundred or
thousand log lines because the earlier log lines for those connections or
mails are in the previous log file; loading the saved state from the
previous log file will solve this problem.} or when there are problems
tracking a child mail (see \sectionref{tracking re-injected mail}).  There
should be few or no warnings when parsing, and when finished parsing the
state table should only contain entries for mails that had yet to be
delivered when the log files ended, or were accepted before the log files
began.

At the time of writing the parser is being tested with \numberOFlogFILES{}
log files.  There are 5 warnings produced, but because the parser errs on
the side of producing more warnings rather than fewer, those 5 warnings
represent 3 instances of 1 problem: 3 connections started before the first
log file, so their initial log lines are missing, leading to warnings when
their remaining log lines are parsed.

@@

    \item 9 mails whose initial log lines were not present in the log
        files.

@@

    \item Queueids and \glspl{pid} naturally group together log lines
        belonging to one mail or connection respectively; it is extremely
        unlikely that a log line would be associated with the wrong
        connection.

@@

It is important that the log lines used are contiguous so that all log
entries are present for as many of the connections as possible.  This log
segment was parsed with all debugging options enabled, resulting in 167,448
lines of output.\footnote{A mean of 27.908 lines of output per line of
input; each connection has 30 debugging lines, plus 21 debugging lines per
result.  Connections which have been cloned will have the cloned connection
in their debugging output, plus another 33 debugging lines.  Those numbers
are approximate, and may vary $\pm{}$ 2.  There is a linear relationship
between the number of log lines and debugging lines: $33(connections) +
30(accepted~~mails) + 21(results)$.  This formula is an approximation only,
and has not been rigorously verified.}  All 167,448 lines were examined in
conjunction with the input log file and a dump of the resulting database,
verifying that for each of the input lines the parser used the correct rule
and executed the correct action, that in turn produced the correct result
and inserted the correct data in the database.  The log file segment
produced 4 warnings, 10 mails remaining in the state tables, and 1625
connections correctly entered in the database.

### 2008-11-22 doc/logparser.tex
Switched to new glossaries package. A couple of other XXX items have been done too.

% Acronyms and glossary entries.
% global=true prevents acronyms being expanded twice
\usepackage[global=true,acronym=true,style=altlist,number=none,toc=true]{glossary}
\makeglossary{}
\makeacronym{}

@@

@@

% The name of the program, so I only have to change it in one place.
\newcommand{\parsername}{\PLP{}}
\newcommand{\parsernames}{\PLP{}'s}

@@

% Pull in the acronyms early, so they can be used throughout the text.
\input{logparser-acronyms.tex}

@@

XXX SWITCH TO NEW GLOSSARY PACKAGE\@.

XXX IS IT POSSIBLE TO HAVE ACRONYMS THAT ARE NOT INCLUDED IN THE ACRONYM
LIST\@?

@@

XXX CHECK THAT I SAY log line AND log file EVERYWHERE\@.

@@

\newpage
\input{logparser-introduction.tex}
\input{logparser-background.tex}
\input{logparser-other-parsers-review.tex}
\input{logparser-architecture.tex}
\input{logparser-implementation.tex}
\input{logparser-results.tex}
\input{logparser-conclusion.tex}

@@

% Redefine the command used to produce the glossary title, because the
% default command produces an unnumbered section whereas I want a numbered
% section.
XXX MOVE POSTFIX DAEMONS INTO A SEPARATE SECTION\@; NEED TO SWITCH TO NEW
GLOSSARY PACKAGE FIRST\@.
\renewcommand{\glossarytitle}{\section{Glossary}\label{Glossary}}
\printglossary{}
% Redefine the command a second time to produce acronyms instead of a
% glossary.
\renewcommand{\glossarytitle}{\section{Acronyms}\label{Acronyms}}
\printacronym{}

\input{logparser-glossary.tex}
% logparser-acronyms.tex has been included in the preamble so that the
% commands are available throughout the text.
\end{document}

>>>>

@@

% Acronyms and glossary entries.
\usepackage[acronym,toc,numberedsection]{glossaries}
\newglossary[plg]{postfix}{pin}{pout}{Postfix Daemons}
\makeglossaries{}

@@

% The name of the program, so I only have to change it in one place.
\newcommand{\parsername}{\gls{PLP}}
\newcommand{\parsernames}{\gls{PLP}'s}

@@

@@

@@

@@

\newpage

% Pull in the acronyms early, so they can be used throughout the text.
\input{logparser-acronyms}
\input{logparser-glossary}
\input{logparser-postfix-daemons}
\glsaddall[types={postfix}]

\input{logparser-introduction}
\input{logparser-background}
\input{logparser-other-parsers-review}
\input{logparser-architecture}
\input{logparser-implementation}
\input{logparser-results}
\input{logparser-conclusion}

@@

% Redefine the command used to produce the glossary title, because the
% default command produces an unnumbered section whereas I want a numbered
% section.
%\renewcommand{\glossarytitle}{\section{Glossary}\label{Glossary}}
\label{Glossary}
\printglossaries{}
% Redefine the command a second time to produce acronyms instead of a
% glossary.
%\renewcommand{\glossarytitle}{\section{Acronyms}\label{Acronyms}}
%\printacronym{}

\end{document}
